---
title: "Forecasting Homework 1"
author: "Group C"
date: "1/23/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages

```{r packages}
library(fBasics)
library(forecast)
library(fUnitRoots)
library(ggplot2)
```

Load Data

```{r load data, echo=FALSE}
hw <- read.csv("/Users/cris/Desktop/Courses/Second Semester/Forecasting/Homework 1/Homework_1_DATA.csv",header=TRUE,sep=";",dec=",")
```

Setting Series

```{r setting series, echo=FALSE}
series1=hw[,1] [1:300]
series2=hw[,2] [1:300] 
series3=hw[,3] [1:300]
series4=hw[,4] [1:300]
series5=hw[,5] [1:2000]
series6=hw[,6]
series7=hw[,7]
```


# Series 1 Analysis

```{r Series 1: Checking for WN, echo=FALSE}
s1 <- series1

par(mfrow=c(3,1))
ts.plot(s1) # mu = 0, mu and var stationary
acf(s1) # close to 0, no lags
pacf(s1) # close to 0, 1 lag in t12

# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test (s1, lag = 20, type="Ljung")
```

Series 1 has a mean equal to 0, is stationary in both the mean and variance, and has an ACF and PACF close to 0. It is WN, next we check for GWN.


```{r Series 1: Checking for GWN}
hist(s1,prob=T,ylim=c(0,0.4),xlim=c(mean(s1)-3*sd(s1),mean(s1)+3*sd(s1)),col="red")
lines(density(s1),lwd=2)
mu<-mean(s1)
sigma<-sd(s1)
x<-seq(mu-3*sigma,mu+3*sigma,length=100)
yy<-dnorm(x,mu,sigma)
lines(x,yy,lwd=2,col="blue")
# formal normality test
	# Ho: the data is normally distributed
	# H1: the data is not normally distributed
shapiro.test(s1) # p-value = 0.01941
```

With a p-value of 0.01941, we REJECT that the data is normally distrubuted, meaning we do not have GWN. Next we check for SWN.

```{r Series 1: Checking for SWN}
par(mfrow=c(3,1))
ts.plot(s1^2)   
acf(s1^2)
pacf(s1^2)
# formal test for white noise (zero autocorrelations)
# Ho: uncorrelated data
# H1: correlated data
Box.test(s1^2,lag=20, type="Ljung") 
```

Series 1 is stationary in the mean, but not in the variance. The data is uncorrelated so we cannot tell if it is SWN.

Series 1 finished.


# Series 2 Analysis

```{r Series 2: Checking for WN}
s2 <- series2

par(mfrow=c(3,1))
ts.plot(s2) # mean != 0, stationary only in variance
acf(s2) # structured, multiple lags
pacf(s2) # unstructed

# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test(s2, lag = 20, type="Ljung")


#Checking for normality
hist(s2,prob=T,ylim=c(0,0.08),xlim=c(mean(s2)-3*sd(s2),mean(s2)+3*sd(s2)),col="red")
lines(density(s2),lwd=2)
mu2<-mean(s2)
sigma2<-sd(s2)
x2<-seq(mu2-3*sigma2,mu2+3*sigma2,length=100)
yy2<-dnorm(x2,mu2,sigma2)
lines(x2,yy2,lwd=2,col="blue")

# shapiro test
shapiro.test(s2)
```

Series 2 does not have a mean equal to 0, is not stationary in the mean, but stationary in the variance. ACF is structued and will require a linear model, PCF is unstructured. Next, we will transform the mean.

```{r Series 2: Transforming the mean}
d=1
z2<-diff(s2,d)

par(mfrow=c(3,1))
ts.plot(z2) # mu = 0, mu and var stationary 
acf(z2) # unstructured
pacf(z2) # unstructured

# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test(z2, lag = 20, type="Ljung")
```

The transformed Series 2 now has a mean equal to 0, stationary in both the mean and the variance, ACF and PCF close to 0. Now we test for GWN.

```{r Series 2: Testing for GWN}
hist(z2,prob=T,ylim=c(0,0.6),xlim=c(mean(z2)-3*sd(z2),mean(z2)+3*sd(y)),col="red")
lines(density(z2),lwd=2)
mu2<-mean(z2)
sigma2<-sd(z2)
x2<-seq(mu2-3*sigma2,mu2+3*sigma2,length=100)
yy2<-dnorm(x2,mu2,sigma2)
lines(x2,yy2,lwd=2,col="blue")
# formal normality test
	# Ho: the data is normally distributed
	# H1: the data is not normally distributed
shapiro.test(z2) # p-value = 0.69
```

With a p-value of .69, we DO NOT reject that Series 2 is normally distributed, making it GWN. We cannot predict.

Series 2 finished.


# Series 3 Analysis

```{r Series 3: Checking for WN}
s3 <- series3

par(mfrow=c(3,1))
ts.plot(s3) # mu !0, mean not stationary, var is stationary
acf(s3) # structured, multiple lags
pacf(s3) # 1 lag

#Checking for normality
hist(s3,prob=T,ylim=c(0,0.05),xlim=c(mean(s3)-3*sd(s3),mean(s3)+3*sd(s3)),col="red")
lines(density(s3),lwd=2)
mu3<-mean(s3)
sigma3<-sd(s3)
x3<-seq(mu3-3*sigma3,mu3+3*sigma3,length=100)
yy3<-dnorm(x3,mu3,sigma3)
lines(x3,yy3,lwd=2,col="blue")

# shapiro test
shapiro.test(s3)
```

Series 3 does not have a mean equal to 0, is not stationary in the mean but is stationary in the variance, and the ACF and PACF are structured. It is not WN and will require a linear model. Next we will transform the mean.

```{r Series 3: Transforming the mean}
d=1
z3 <- diff(s3,d)

par(mfrow=c(3,1))
ts.plot(z3) # mu = 0, mu and var stationary 
acf(z3) # structured
pacf(z3) # unstructured

#Checking for normality
hist(z3,prob=T,ylim=c(0,0.5),xlim=c(mean(z3)-3*sd(z3),mean(z3)+3*sd(z3)),col="red")
lines(density(z3),lwd=2)
mu3z<-mean(z3)
sigma3z<-sd(z3)
x3z<-seq(mu3z-3*sigma3z,mu3z+3*sigma3z,length=100)
yy3z<-dnorm(x3z,mu3z,sigma3z)
lines(x3z,yy3z,lwd=2,col="blue")

# shapiro test
shapiro.test(z3)
```

The transformed Series 3 has a mean equal to 0, is stationary in both the mean but not in the variance. The ACF is structured and PACF is unstructued. It is not WN and will need a linear model.

Series 3 finished.


# Series 4 Analysis

```{r Series 4: Checking for WN}
s4 <- series4

par(mfrow=c(3,1))
ts.plot(s4) # mu != 0, mu and var not stationary
acf(s4) # structured
pacf(s4) # not structured

#Checking for normality
hist(s4,prob=T,ylim=c(0,0.0005),xlim=c(mean(s4)-3*sd(s4),mean(s4)+3*sd(s4)),col="red")
lines(density(s4),lwd=2)
mu4<-mean(s4)
sigma4<-sd(s4)
x4<-seq(mu4-3*sigma4,mu4+3*sigma4,length=100)
yy4<-dnorm(x4,mu4,sigma4)
lines(x4,yy4,lwd=2,col="blue")

# shapiro test
shapiro.test(s4)
```

Series 4 does not have a mean equal to 0, is not stationary in both the mean and variance. The ACF is structured and the PACF is unstructured. We will need a linear model and will need to transform the mean.

```{r Series 4: Transforming the mean}
d=1
z4 <- diff(s4,d)

par(mfrow=c(3,1))
ts.plot(z4) # mu !0, mu and var not stationary
acf(z4) # structured
pacf(z4) # structured 

#Checking for normality
hist(z4,prob=T,ylim=c(0,0.05),xlim=c(mean(z4)-3*sd(z4),mean(z4)+3*sd(z4)),col="red")
lines(density(z4),lwd=2)
mu4z<-mean(z4)
sigma4z<-sd(z4)
x4z<-seq(mu4-3*sigma4,mu4+3*sigma4,length=100)
yy4z<-dnorm(x4z,mu4z,sigma4z)
lines(x4z,yy4z,lwd=2,col="blue")

# shapiro test
shapiro.test(z4)
```

The transformed Series 4 still needs to be transformed.

```{r Series 4: Transforming the mean again}
d=1
z4z <- diff(z4,d)

par(mfrow=c(3,1))
ts.plot(z4z) # mu = 0, mu and var are stationary
acf(z4z) # structured
pacf(z4z) # unstructued

# shapiro test
shapiro.test(z4z)
```

The newly transformed Series 4 now has a mean equal to 0, is stationary in both the mean and variance. The ACF is structured and the PACF is unstructued. It will require a linear model.

Series 4 finished.


# Series 5 Analysis

```{r Series 5: Checking for WN}
s5 <- series5

par(mfrow=c(3,1))
ts.plot(s5) # mu = 0, mu stationary, var not stationary
acf(s5) # unstructured
pacf(s5) # unstructued

# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test(s5, lag = 20, type="Ljung")
```

Series 5 has a mean equal to 0, is stationary in the mean but not in the variance. ACF and PACF are unstructured. With a p-value of 0.03248, we REJECT that it is WN.

Series 5 finished.


# Series 6 Analysis

```{r Series 6: Checking for WN}
s6 <- series6

par(mfrow=c(3,1))
ts.plot(s6) # mu = 0, mu stationary, var not stationary
acf(s6) # structured
pacf(s6) # not structued

# testing for normality
hist(s6,prob=T,ylim=c(0,0.7),xlim=c(mean(s6)-3*sd(s6),mean(s6)+3*sd(s6)),col="red")
lines(density(s6),lwd=2)
mu6 <- mean(s6)
sigma6 <- sd(s6)
x6 <- seq(mu6-3*sigma6,mu6+3*sigma6,length=100)
yy6 <- dnorm(x6,mu6,sigma6)
lines(x6,yy6,lwd=2,col="blue")
# formal normality test
	# Ho: the data is normally distributed
	# H1: the data is not normally distributed
shapiro.test(s6)
```

Series 6 the mean is equal to 0, the mean is stationary but the variance is not. The ACF and PCF are structued, so it will require a linear model.

Series 6 finished.


# Series 7 Analysis

```{r Series 7: Checking for WN}
s7 <- series7

par(mfrow=c(3,1))
ts.plot(s7) # mu != 0, mu not stationary, var stationary
acf(s7) # structured
pacf(s7) # structured

# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test(s7, lag = 20, type="Ljung")

# testing for normality
hist(s7,prob=T,ylim=c(0,0.05),xlim=c(mean(s7)-3*sd(s7),mean(s7)+3*sd(s7)),col="red")
lines(density(s7),lwd=2)
mu7 <- mean(s7)
sigma7 <- sd(s7)
x7 <- seq(mu7-3*sigma7,mu7+3*sigma7,length=100)
yy7 <- dnorm(x7,mu7,sigma7)
lines(x7,yy7,lwd=2,col="blue")
# formal normality test
	# Ho: the data is normally distributed
	# H1: the data is not normally distributed
shapiro.test(s7) # p-value < 2.2e-16
```

Series 7 does not have a mean equal to 0, its mean is not stationary but its variance is. The ACF and PACF are structured so a linear model will be needed. Next, we will need to transform the mean.

```{r Series 7: Transforming the Mean}
d = 1
z7 = diff(s7,d)
par(mfrow = c(3,1))
ts.plot(z7) # mu = 0, mu stationary, var not stationary
acf(z7) # not structured
pacf(z7) # not structured


hist(z7,prob=T,ylim=c(0,.7),xlim=c(mean(z7)-3*sd(z7),mean(z7)+3*sd(z7)),col="red")
lines(density(z7),lwd=2)
mu7z <- mean(z7)
sigma7z <- sd(z7)
x7z <- seq(mu7z-3*sigma7z,mu7z+3*sigma7z,length=100)
yy7z <- dnorm(x7z,mu7z,sigma7z)
lines(x7z,yy7z,lwd=2,col="blue")
# Ho: uncorrelated data - white noise
# H1: correlated data
Box.test(z7, lag = 20, type="Ljung")
```

The transformed Series 7 has a mean equal to 0, it is stationary in the mean but not in the variance. The ACF and PCF are not structured, so it is WN. Next, we test for GWN.

```{r Series 7: Testing for GWN}
hist(z7,prob=T,ylim=c(0,0.7),xlim=c(mean(z7)-3*sd(z7),mean(z7)+3*sd(z7)),col="red")
lines(density(z7),lwd=2)
mu7 <- mean(z7)
sigma7 <- sd(z7)
x7 <- seq(mu7-3*sigma7,mu7+3*sigma5,length=100)
yy7 <- dnorm(x7,mu7,sigma7)
lines(x7,yy7,lwd=2,col="blue")
# formal normality test
	# Ho: the data is normally distributed
	# H1: the data is not normally distributed
shapiro.test(z7) # p-value < 2.2e-16
```

With a p-value less than 2.2e-16, we can REJECT that the data is normally distributed. We now test for SWN.

```{r Series 7: Checking for SWN}
par(mfrow=c(3,1))
ts.plot(z7^2) # mu = 0, stationary in mean but not variance  
acf(z7^2) # structured
pacf(z7^2) # structured
# formal test for white noise (zero autocorrelations)
# Ho: uncorrelated data
# H1: correlated data
Box.test(z7^2,lag=20, type="Ljung")    # Null: ro1=…=ro20=0
```

1It is stationary in the mean but not in the variance, but there is structure in the ACF and PCF, so it is not SWN. We will need a non-linear model.

Series 7 finished.